---
title: "Lab1_AF"
author: "Hannah Wasson"
date: "2024-06-25"
output: html_document
---

#Installing and using normtemp package to answer questions... 

```{r}
#install.packages('UsingR')
library(UsingR)
data(normtemp)
```

Use the normtemp dataset to answer the following:
a. Determine the following statistics for the variable temperature:
b. Does temperature appear to be normally distributed? no. 

```{r}
library(ggplot2)

min(normtemp$temperature)
max(normtemp$temperature)
mean(normtemp$temperature)
sd(normtemp$temperature)
ggplot(data = normtemp, aes(x = temperature)) + geom_histogram()
```

c. Create box plots for temperature. Are there any outliers? Display
a reference line at 98.6
Does the median body temperature seem to be 98.6 degrees? yes. 


```{r}
ggplot(data = normtemp, aes(x = temperature)) + geom_boxplot()+ geom_vline(xintercept=98.6) 
```
Using the Ameshousing dataset from our in-class examples, run some
distributional analysis on Sale_Price, Log(Sale_Price), and Gr_Liv_Area.
a. Create histograms of these three variables.
- Overlay a kernel density estimator of the variables.

```{r}
library(AmesHousing)
ames <- make_ordinal_ames()

ggplot(data = ames, aes(x = Sale_Price)) + geom_histogram(aes(y=..density..), fill = 'lightblue') + geom_density()

ggplot(data = ames, aes(x = log(Sale_Price))) + geom_histogram(aes(y=..density..), fill = 'orange') + geom_density()

ggplot(data = ames, aes(x = Gr_Liv_Area)) + geom_histogram(aes(y=..density..), fill = 'lightgreen') + geom_density()



```


b. Create a QQ Plot for both Sale_Price and Log(Sale_Price). Based on
these exploratory procedures, which version of the price information
(Sale_Price or Log(Sale_Price)) would you say is closer to being normally
distributed?

```{r}
ggplot(data = ames, aes(sample = Sale_Price), color = 'lightblue') + geom_qq() + stat_qq() + stat_qq_line() + labs(title = 'Sales Price QQ Plot')
ggplot(data = ames, aes(sample = log(Sale_Price)), color = 'orange') + geom_qq() + stat_qq() + stat_qq_line()+ labs(title = 'log(Sales Price) QQ Plot')
ggplot(data = ames, aes(sample = Gr_Liv_Area), color = 'lightgreen') + geom_qq() + stat_qq() + stat_qq_line()+ labs(title = 'Gr_Liv_Area QQ Plot')
```

log(Sale_Price) is the closest to being normally distributed. 


# Using the Ameshousing dataset from our in-class examples, determine the
# following:

a. What type of variables are each of these columns (Nominal, Ordinal, or
Continuous/Quantitative)? Keep in mind that the way they are
represented in the R dataset may not be appropriate, so you should make
this determination using your own judgement based on the data you are
looking at.
- Overall_Qual -- Ordinal 
- Lot_Shape --  Ordinal
- Heating_QC -- Ordinal
- Lot_Area -- Quantitative


```{r}
str(ames$Overall_Qual)
str(ames$Lot_Shape)
str(ames$Heating_QC)
str(ames$Lot_Area)

```

