---
title: "AF_Lab_2"
author: "Hannah Wasson"
date: "2024-06-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Revisit the NormTemp dataset from Lab 1, where we examined the observed mean body temperature (temperature) in comparison to the well-known “average” of 98.6

Perform a statistical test (alpha = 0.05) to determine whether this well- known number is actually the mean body temperature. What is your p-value? Explain in words what this p-value means. What is your conclusion?

```{r}

library(UsingR)
data(normtemp)

t.test(normtemp$temperature, mu = 98.6)
```

The p-value is very low so we will reject our null hypothesis and concluded the the well-know  body temperature is statistically different than the observed body temperature.  


# Give the 95% Confidence Interval for temperature. Explain in words what a 95% confidence interval represents.

```{r}
t.test(normtemp$temperature, mu = 98.6, conf.level = 0.95)

```

A 95 % confidence interval represents the range of values within which you are 95% confident. 

# If we restrict our analysis to only the females in this dataset, would our conclusion change?

To do this we will need to split our dataset by gender. 

```{r}
t.test(normtemp$temperature[normtemp$gender == 2], mu = 98.6, conf.level = 0.95)
```

No, the conclusion does not change. We still would still reject the null hypothesis at an alpha level of 0.05


# Is there any difference (alpha=0.05) in temperature between the two genders recorded in this dataset (be sure to look at assumptions and perform the correct test)?

First we need to check for normality. We can check this by looking at the QQ-plot or the Shapiro test. Let's do both. 
```{r}
# males 
shapiro.test(normtemp$temperature[normtemp$gender == '1'])

# females 
shapiro.test(normtemp$temperature[normtemp$gender == '2'])

library(ggplot2)

ggplot(data = normtemp, aes(sample = temperature, color = factor(gender))) + geom_qq() + geom_qq_line()

```


By the looks of the QQ-Plot and the Shapiro Test we can assume normality. Let's look at variance now. 

```{r}
var.test(temperature ~ gender, data = normtemp)
```

With an alpha value of 0.05, we conclude that variances are not equal and fail to reject the null. Therefore, in my t-test I will assign variances to be not equal to each other and use Satterhwaite's t-test. 

```{r}
t.test(temperature ~ gender, data = normtemp, var.equal = FALSE)
```

Our p-value is less than alpha of 0.05 so we will reject the null hypothesis and conclude that there is statistically a significant difference between female and male mean body temperature. We CANNOT say that one is greater than the other, only that there is a difference. 


# The Airline dataset contains information regarding the number of international airline travelers (variable air) across different months of the year from 1949-1960. To obtain this data set, you will need to:

We are interested in knowing if during this time period there was a significant difference between air travel in the Summer months of June, July, and August vs. the remainder of the year? Use a statistical hypothesis test (alpha=0.05) to support your answer. In order to get month information, you will need to:

```{r}
library(tseries)
library(forecast)
library(dplyr)
cycle(AirPassengers)

air1 = data.frame(AirPassengers)
air2 = air1 %>% mutate(summer=ifelse(cycle(AirPassengers) %in%
6:8,1,0))

```

First we need to check normality and variances. 

```{r}
# summer 
shapiro.test(air2$AirPassengers[air2$summer == '1'])

# not summer 
shapiro.test(air2$AirPassengers[air2$summer == '0'])

library(ggplot2)

ggplot(data = air2, aes(sample = AirPassengers, color = factor(summer))) + geom_qq() + geom_qq_line()

ggplot(data = air2, aes(x = AirPassengers, color = factor(summer))) + geom_histogram()

var.test(AirPassengers ~ summer, data = air2)


```
Cannot verify normality and the variances are not equal. We must go with a non-parametric test. This is the Wilcoxon Rank test ...

```{r}
wilcox.test(AirPassengers ~ summer, data = air2)
```

The p-value is less than 0.05 so we can reject the null hypothesis. We can conclude that there is a difference in location and distributional dominance. 



