---
title: "af_lab_9"
author: "Hannah Wasson"
date: "2024-07-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Diagnostic Testing 

## 1. Using the cars2010 data set, run the regression with the following explanatory variables: EngDispl, Transmission, AirAspirationMethod, TransLockup, TransCreeperGear, DriveDesc, IntakeValvePerCyl, CarlineClassDesc, VarValveLift

```{r}
library(ggplot2)
library(tidyverse)


```


### a. Use plots to identify potential influential observations based on the suggested cutoff values.

Using DFFITS ...

```{r}
n.index=seq(1,nrow(cars)) 
cars=cbind(cars,n.index)

lm_full <- lm(FE ~ EngDispl + Transmission + AirAspirationMethod + TransLockup + TransCreeperGear + DriveDesc + IntakeValvePerCyl + CarlineClassDesc + VarValveLift, data = cars)

resid_full_mod <- tibble(residuals(lm_full))
resid_full_mod %>% arrange(desc(resid_full_mod))


length(coefficients(lm_full))

df.cut=2*(sqrt((length(coefficients(lm_full)))/nrow(cars)))

ggplot(lm_full,aes(x=n.index,y=dffits(lm_full))) + geom_point(color="orange")+geom_line(y=df.cut)+geom_line(y=-df.cut)+labs(title = "DFFITS",x="Observation",y="DFFITS")


```

### b. Are there any observations with a dffits larger than 1 AND studentized residuals larger than 3 in magnitude? If so, list the observations.

```{r}
dffits_tib <- tibble(dffits = dffits(lm_full))
dffits_tib %>%  filter(dffits > 1) %>% arrange(desc(dffits))

```


Studentized Residuals:

```{r}

hat.cut=2*(length(coefficients(lm_full)))/nrow(cars)

ggplot(lm_full,aes(x=n.index,y =hatvalues(lm_full)))+geom_point(color="blue")+geom_line(y=hat.cut)+labs(title = "Hat values",x="Observation",y="Hat Values") 

ggplot(lm_full,aes(x=hatvalues(lm_full),y=rstudent(lm_full)))+geom_point(color="orange")+geom_line(y=-3)+geom_line(y=3)+labs(title = "External Studentized Residuals",x="Hat values",y="Residuals") + geom_vline(xintercept = hat.cut)

student_tib <- tibble(student_res = rstudent(lm_full))
student_gr <- student_tib %>%  filter(student_res > 3) %>% arrange(desc(student_res)) %>% count()
student_ls <- student_tib %>%  filter(student_res < -3) %>% arrange((student_res)) %>% count()

ggplot(lm.model,aes(x=n.index,y=rstudent(lm.mod
el)))+geom_point(color="orange")+geom_line(y=-
3)+geom_line(y=3)+labs(title = "External
Studentized
Residuals",x="Observation",y="Residuals")

print(student_gr + student_ls)
```


```{r}
cooks.cut=4/(nrow(cars2010)-length(coefficients(lm_full))-1)
ggplot(lm_full,aes(x=n.index,y=cooks.distance(lm_full)))+geom_point(color="orange")+geom_line(y=cooks.cut)+labs(title = "Cook's D",x="Observation",y="Cook's Distance")

cooks_tib <- tibble(cooks_tib_val = cooks.distance(lm_full))
cooks_tib %>% arrange(desc(cooks_tib_val)) %>% signif(digits=3)
```

